An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-11-29 paper/1.introduction.tex

The composition of BX programs is a fundamental construct to build more complex BX programs. 
%In BX programs, composition of programs is essential for writing various interesting evaluations.
Because recursions are achieved by compositions, large number of programs include compositions.
In the BX lens papers \cite{} compositions are also known as central part of BX programs.
In this section we use a small BX program that contains compositions as a running example, because the programs including recursions can be complicated easily. The running examples are two versions of two composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).

@@

If the programs are simple like these running examples, we can transform left associative compositions to right associative. However, this is not always possible. \textcolor{red}{For example, let us consider the following 'breverse' program.}

%$\tab
%bsnoc = Rearr\\
%bfoldr bx = Case \ f_1 \ f_2 \ bx_s \text{where: } f_1 = , f_2% = \\

%breverse = RearrS \ f_1 \ f_2 \ bx_s \ \text{where: } f_1 = \l%ambda s.(s.()), \, f_2 = \lambda (s.()).s,\\
%\qtab bx_s = bfoldr bsnoc$

\input{breverse.tex}

\noindent \textcolor{red}{In this program, recursion occurs on the left-hand side of composition. Because of priority, it is impossible to tranform this program to right associative compositions. At the same time, we can not apply fusion \cite{} to this program because this program is dynamically produced depending on the length of the input list.}
Therefore it is important to have a fast evaluation method for left associative compositions.

>>>>

The composition of BX programs is a fundamental construct to build more complex BX programs. 
%In BX programs, composition of programs is essential for writing various interesting evaluations.
Because recursions are achieved by compositions, large number of programs include compositions.
In the BX lens papers \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487} compositions are also known as central part of BX programs.
In this section we use a small BX program that contains compositions as a running example, because the programs including recursions can be complicated easily. The running examples are two versions of two composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).

@@

\textcolor{red}{If the programs are simple like the running examples, we can transform left associative compositions to right associative. However, this is not always possible. For example, let us consider a BX 'reverse' program, $breverse$. $breverse$ is constructed by application of $bfoldr$ (the actual program is shown in Section 2) and $bsnoc$. 
A part of the definition of $bfoldr$, we have $(\product{Replace}{(bfoldr \ bx)}) \circ bx$. In this program, recursion occurs on the left-hand side of composition. Because of priority, it is impossible to transform this program to right associative compositions. At the same time, we can not apply program fusion \cite{Wadler:1988:DTP:80099.80104} to this program because this BX program is dynamically produced depending on the length of the input list. Therefore, it is important to have a fast evaluation method for left associative compositions.}
